Certified Symbolic Finite Transducers: Formalization and Applications to String Analysis

### 2025-09-10 Paper/abstract.tex
Continuing with intro
fix motivation

\begin{abstract}
    Finite Transducers (FTs) extend the capabilities of Finite Automata (FAs) by
    enabling the transformation of input strings into output strings. In many
    practical applications --- including program analysis, string constraint
    solving, and analysis of security-critical sanitizers --- Symbolic FTs 
    (SFTs) and Symbolic FAs (SFAs) are used instead of the explicitly
    represented models. To circumvent to the notorious state-space explosion 
    problem caused by the extremely large alphabet size (e.g. UTF-16), SFTs
    and SFAs allow the representation of the alphabet as an effective boolean
    algebra including finite unions of intervals, as well as 
    SMT-Algebras.
    %can be instantiated with various Boolean algebras, 
    The
    security-critical nature of many of these applications demands trustworthy
    implementations of such systems. To this end, we present the first 
    formalization of SFTs and their most important algorithms in
    Isabelle/HOL. 
    %This symbolic approach enables not only efficient FT operations—such as computing the output language of an FT given a regular language—but also supports infinite alphabets.
%
    To evaluate the effectiveness of our formalization, we apply the formalized 
    SFTs to two applications: (1) \texttt{HTMLdecode}, a sanitizer used for 
    preventing XSS attacks, and (2) string solving, which increasingly
    employs intricate string replacement operations.
    %a string solver that models replacement operations with various semantics, including left-most and replace-all matching. 
    Our experimental results demonstrate that our methods are competitive with
    the existing unverified implementations.

>>>>

\begin{abstract}
    Finite Transducers (FTs) extend the capabilities of Finite Automata (FAs) by
    enabling the transformation of input strings into output strings. In many
    practical applications --- including program analysis, string constraint
    solving, and analysis of security-critical sanitizers --- Symbolic FTs 
    (SFTs) and Symbolic FAs (SFAs) are used instead of the explicitly
    represented models. To circumvent to the notorious state-space explosion 
    problem caused by an extremely large alphabet size (e.g. UTF-16), SFTs
    and SFAs allow the representation of the alphabet as an effective boolean
    algebra including finite unions of intervals, as well as 
    SMT-Algebras.
    %can be instantiated with various Boolean algebras, 
    The
    security-critical nature of many of these applications demands trustworthy
    implementations of such systems. To this end, we present the first 
    formalization of SFTs and their most important algorithms in
    Isabelle/HOL. 
    %This symbolic approach enables not only efficient FT operations—such as computing the output language of an FT given a regular language—but also supports infinite alphabets.
%
    To evaluate the effectiveness of our formalization, we apply the formalized 
    SFTs to two applications: (1) sanitizers for web applications used for 
    preventing XSS attacks, and (2) string solving, which increasingly
    employs intricate string replacement operations.
    %a string solver that models replacement operations with various semantics, including left-most and replace-all matching. 
    Our experimental results demonstrate that our methods are competitive with
    the existing unverified implementations.

### 2025-09-10 Paper/introduction.tex
fix motivation

To evaluate the effectiveness and efficiency of our formalization, we 
provide two applications: (1) analysis of the sanitizer \texttt{HTMLdecode},
which is used for preventing XSS attacks, and (2) an SMT-LIB \cite{smtlib}
compliant certified string solver handling equality, concatenation, regular
constraints, and more important the string replaceall operator, which has been
argued to be indispensable for analysis of XSS 
\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,uss/HooimeijerLMSV11}).
The experimental results demonstrate that our formalization achieves computational efficiency in constraint-solving scenarios.

>>>>

To evaluate the effectiveness and efficiency of our formalization, we 
provide two applications: (1) analysis of the web application sanitizers for html, css, json, and javascript
which are used for preventing XSS attacks, and (2) an SMT-LIB \cite{smtlib}
compliant certified string solver handling equality, concatenation, regular
constraints, and more importantly the string replaceall operator, which has been
argued to be indispensable for analysis of XSS 
\cite{DBLP:conf/popl/LinB16,Kern,Berkeley-JavaScript,DBLP:conf/sp/SaxenaAHMMS10,uss/HooimeijerLMSV11}).
The experimental results demonstrate that our formalization achieves computational efficiency in constraint-solving scenarios.
